---
title: "Social Network Analysis"
author: "Pedro Antonio Benito Rojano"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Social Network Analysis

## Libraries

Library for R Markdown.

```{r}
library(rmarkdown)
```

Library for data frames processing.

```{r message=FALSE}
library(dplyr)
```

Library for read dataset.

```{r}
library(readr)
```

Library for graphs.

```{r message=FALSE}
library(igraph)
```

## Load dataset

Set paths.

```{r}
path <- "../virusTotal/data/clustering.csv"
```

```{r message=FALSE}
virusTotal <- read_csv(path)
```

## Preprocessing

### View dataset

```{r}
virusTotal %>% paged_table()
```

### Remove extra row number columns

As can be seen, at read the col ...1 is added, but the column n already has this information.

```{r}
df <- 
  virusTotal %>% 
  select(-...1)
```

## From data frame to graph

The objective is to generate a graph from existing data frame. But how? 

The idea is create relationships between virus which has same values. Also give weights to relationships as a function of how likely it is that coincidence will occur for that variable.

Plan:

1. Find virus with sames values in any variable.
2. Compute how likely it is that coincidence.
3. Set the Weight of that relationship based on that probability.

What is needed?

1. Create a table with the two first columns with the indexes of the viruses in the relationship. The rest of columns are the values that has in common. 
2. Compute a list with the frequency of each element for every column.
3. Add the column Weight to the data frame based in a frecuency ponderation of each value in common.
4. Create a graph from it.

### Find matching values

```{r}
pairs <- 
  function(n){
    pairs <- list()
    for(i in 1:(n-1)){
      for(j in (i+1):n){
        pairs[[length(pairs) + 1]] <- 
          c(i, j)
      }
    }
    names(pairs) <- 
      pairs %>% 
      sapply(function(v) paste(as.character(v), collapse = "-"))
    return(pairs)
  }
```

```{r}
df_ij <- 
  function(df, i, j){
    df_ij <- 
      (df[i,] == df[j,]) %>% 
      ifelse(as.vector(df[i,]), NA) %>% 
      unlist() %>% 
      c(i, j, .)
    names(df_ij) <- 
      c("v1", "v2", colnames(df))
    return(df_ij)
  }
```

```{r eval=FALSE}
df_edges <- 
  df[1:10] %>%
  ncol() %>% 
  pairs() %>% 
  sapply(function(v) df_ij(df[1:2,], v[1], v[2])) %>% 
  rbind()
```

```{r eval=FALSE}
df_edges %>% 
  paged_table()
```

Not endded.












