---
title: "Virustotal_aRules"
author: "Ángel Sergio"
date: '2022-05-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r librerias}
library(dplyr)
library(arules)
library(magrittr)
library(rmarkdown)
```


```{r lectura, include=FALSE}
virus <- read.csv('virusTotal.csv')
virus$additional_info.androguard.RiskIndicator.PERM.PRIVACY

```

```{r lectura, include=FALSE}
rules <- virus %>%
   select(-c(additional_info.androguard.RiskIndicator.PERM.DANGEROUS,
             additional_info.androguard.RiskIndicator.PERM.GPS,
             additional_info.androguard.RiskIndicator.PERM.NORMAL,
             additional_info.androguard.RiskIndicator.PERM.PRIVACY
             
             
             ))



```

## R Markdown


# Puntuación global
futbin$rating <- ordered(cut(futbin$rating,
                            unique(quantile(futbin$rating)),
                            include.lowest = TRUE))
# Velocidad (análogo para disparo, pase, regate, defensa y físico)
futbin$pac <- ordered(cut(futbin$pac,
                          c(0, 65, 75, 80, 100), 
                          labels = c("Muy bajo", "Bajo", "Alto", "Muy alto"),
                          include.lowest = TRUE))
# Altura
futbin$hei <- ordered(cut(futbin$hei, c(0, 177, 185, 205),
                          labels = c("Bajo", "Alto", "Muy alto"),
                          include.lowest = TRUE))
# Popularidad
futbin$popularity <- ordered(cut(futbin$popularity,
                                 quantile(futbin$popularity),
                                 include.lowest = TRUE))
# Precio
futbin$ps <- ordered(cut(futbin$ps,
                         quantile(futbin$ps),
                         include.lowest = TRUE))
# Pie malo
futbin$weak_foot <- factor(futbin$weak_foot,
                          levels = c("1", "2", "3", "4", "5"),
                          ordered = TRUE)
# Filigranas
futbin$skills    <- factor(futbin$skills,
                          levels = c("1", "2", "3", "4", "5"),
                          ordered = TRUE)