---
title: "Clustering"
author: "Pedro Antonio Benito Rojano"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Clustering

## Libraries

Library for R Markdown.

```{r}
library(rmarkdown)
```

Library for clustering

```{r message=FALSE}
library(factoextra)
```

Library for data frames processing.

```{r message=FALSE}
library(dplyr)
```

Library for read dataset.

```{r}
library(readr)
```

Library for data presentation.

```{r message=FALSE}
library(scales)
```

Library for manage strings.

```{r}
library(stringr)
```

## Configuration

### Seed

Set seed.

```{r}
set.seed(0)
```

### Paths

Set paths.

```{r}
path_import <- "data/preprocessed.csv"
path_export <- "data/clustering.csv"
```


## Load dataset

```{r message=FALSE}
virusTotal <- read_csv(path_import)
```
## Preprocessing

### View dataset

```{r}
virusTotal %>% paged_table()
```

### Remove row number columns

As can be seen, at read the col ...1 is added, but the column n already has this information.

Both of them are not interesting columns for clustering, so remove them.

```{r}
df <- 
  virusTotal %>% 
  select(-...1, -n)
```

### Cast logical columns as numeric

Cast logical to numeric in order to be able to cluster with them.

```{r}
df <- 
  df %>% 
  mutate_if(is.logical, as.numeric)
```

### Select numeric columns

Only numeric columns are suitable for clustering, so just remove them.

```{r}
df <- 
  df %>% 
  select_if(is.numeric)
```

### Rows with NA

```{r}
rows_with_na <- 
  function(df){
      df %>% 
      is.na() %>% 
      rowSums() %>% 
      sapply(function(x) x > 0) %>% 
      which()
  }
```

```{r}
nrows_with_na <- 
  df %>% 
  rows_with_na() %>% 
  length()

nrows_with_na / nrow(df)
```

Most of the rows of the dataframe has NA so can't be deleted. If rows with NA were deleted most of the information will be loss.

### Replace NA

Clustering do not work with NAs because distance measure is needed. So just replace NA for predicted value using lineal model.

```{r}
remove_0 <- 
  function(x) x[x!=0]
```

```{r}
num_of_NA_by_column <- 
  function(df){
    df %>% is.na() %>% colSums()
  }
```

```{r}
names_of_colums_with_NA <- 
  function(df)
    df %>% 
      num_of_NA_by_column() %>% 
      remove_0 %>% 
      names()
```

```{r}
lm_from_all_columns <- 
  function(df, column){
    target_column_name <- 
      colnames(df)[column]
    col <- df[,column]
    df <- 
      df %>% 
      select(., -all_of(names_of_colums_with_NA(.))) %>% 
      cbind(col)
    formula_lm <- 
      paste(target_column_name, ".", sep = " ~ ") %>% 
      as.formula()
    lm(formula_lm, data = df, na.action = na.omit)
  }
```

```{r}
replace_NA_by_predicted_col <- 
  function(df, col_index){
    col <- 
      df[, col_index] %>% 
      unlist()
    has_na <- 
      col %>% 
      anyNA()
    if(has_na){
      lm_col <- 
        lm_from_all_columns(df, col_index)
      col_predicted <- 
        predict.lm(lm_col, df) %>% 
        as.double()
      na_col <- 
        col %>% 
        sapply(is.na) %>% 
        as.vector()
      col <- 
        if_else(na_col, col_predicted, col)
    }
    col <- 
      as.vector(col)
    return(col)
  }
```

```{r}
replace_NA_by_predicted <- 
  function(df){
    for(index in 1:ncol(df)){
      df[,index] <- replace_NA_by_predicted_col(df, index)
    }
    return(df)
  }
```

```{r warning=FALSE}
df <- 
  replace_NA_by_predicted(df)
```

### View result

```{r}
df %>% 
  paged_table()
```

## Clustering

### K-Means

#### Optimal number of clusters

Plot a graphic about the optimal number of clusters. 

```{r}
df %>% 
  scale() %>% 
  fviz_nbclust(kmeans, method = "wss", k.max = 15)
```

As can be seen 3, 7 and 10 clusters seem to be optimal number of clusters.

#### Define a function

```{r}
cluster_k <- function(df, k){
  df_scale <- 
    df %>% 
    scale() %>% 
    as.data.frame()
  l <- list()
  l$kmeans <- 
    kmeans(df_scale, centers = k, nstart = 25)
  l$plot <-
    l$kmeans %>% 
    fviz_cluster(geom = "point", data = df_scale) + 
    ggtitle(paste("k", k, sep="="))
  l$table <- tibble(row = 1:nrow(df), cluster = l$kmeans$cluster, df)
  l$cluster_size <- 
    l$table %>% 
    group_by(cluster) %>% 
    summarise(size=n())
  return(l)
}
```

#### K-3

Apply kmeans with 3 clusters.

```{r}
cluster_k3 <- 
  cluster_k(df, 3)
```

Show results

```{r}
cluster_k3$plot
cluster_k3$table %>% 
  paged_table()
cluster_k3$cluster_size %>% 
  paged_table()
```

As can be seen in the table of the size of the clusters there are two main groups, but the other has only two virus in the cluster 2.

#### K-7

Apply kmeans with 7 clusters.

```{r}
cluster_k7 <- 
  cluster_k(df, 7)
```

Show results

```{r}
cluster_k7$plot
cluster_k7$table %>% 
  paged_table()
cluster_k7$cluster_size %>% 
  paged_table()
```

Again in k-7 clusters there are four main groups but the others has only 1 or 2 viruses.

That is so curious.

#### K-10

Apply kmeans with 10 clusters.

```{r}
cluster_k10 <- 
  cluster_k(df, 10)
```

Show results

```{r}
cluster_k10$plot
cluster_k10$table %>% 
  paged_table()
cluster_k10$cluster_size %>% 
  paged_table()
```

Again in k-10 clusters there are many small groups that are very different from the rest in terms of number of viruses clasified in.

#### Conclusion

This is so strange clustering distribution, so would be interesting to inspect why it is happening.

#### Add cluster information to the dataframe.

```{r}
df_clustering_info_added <- 
  df %>% 
  mutate(
    analysis_info.clustering.cluster.k3 = 
      cluster_k3$kmeans$cluster) %>% 
  mutate(
    analysis_info.clustering.cluster.k7 = 
      cluster_k7$kmeans$cluster) %>% 
  mutate(
    analysis_info.clustering.cluster.k10 = 
      cluster_k10$kmeans$cluster)
```


### Inspect small groups 

Define a function.

```{r}
small_clusters <- 
  function(table_cluster_size, percent){
    max <- 
      max(table_cluster_size$size)
    min <- 
      min(table_cluster_size$size)
    threshold <- 
      min + (max - min) * percent
    small_clusters <- 
      table_cluster_size %>% 
      filter(size <= threshold) %>% 
      select(cluster) %>% 
      unlist()
    return(small_clusters)
  }
```

```{r}
belongs_to_small_cluster <- 
  function(cluster, small_clusters)
    cluster %in% small_clusters
```

#### Small clusters of k10

Compute small clusters of k10

```{r}
small_clusters_k10 <- 
  cluster_k10$cluster_size %>% 
  small_clusters(0.2)
```

Show rows that belong to small clusters.

```{r}
belongs_to_small_cluster_k10 <- 
  belongs_to_small_cluster(
    cluster_k10$table$cluster,
    small_clusters_k10
  )

small_groups_df <- 
  df %>% 
  filter(belongs_to_small_cluster_k10)
small_groups_df %>% 
  paged_table()
```
#### Add this information to the dataframe.

```{r}
df_clustering_info_added <- 
  df_clustering_info_added %>% 
  mutate(
    analysis_info.clustering.belongs_to_small_cluster.k10 =
      belongs_to_small_cluster_k10)
```

### Dendogram

Calculate distances.

```{r}
distances <- 
  df %>% 
  scale() %>% 
  get_dist(method="euclidean")
```

Generate dendogram.

```{r}
dendogram <- 
  distances %>% 
  hclust(method="ward.D")
```

Plot dendogram.

```{r}
dendogram %>% 
  plot(main="Dendogram")
```
### Analyse clusters

#### Analyse variables by clusters

The idea is view how affect belonging to a cluster to the main variables distribution.

The main variables are those whose names do not begin with "something.". So just keep columns whose name do not have '.'.

```{r}
cols_to_select <- 
  df_clustering_info_added %>% 
  colnames() %>% 
  str_detect("\\.", negate = TRUE)
```

```{r}
means_by_cluster <- 
  df_clustering_info_added %>% 
  group_by(analysis_info.clustering.cluster.k10) %>% 
  select_if(cols_to_select) %>% 
  summarise_all(mean) %>% 
  round(2) %>% 
  rename_with(function(name) paste("mean", name, sep = "_")) %>% 
  rename(cluster = mean_analysis_info.clustering.cluster.k10) %>% 
  arrange(desc(mean_size))
means_by_cluster %>% 
  paged_table()
```

#### Analyse variables by cluster size

The idea is view how affect the cluster size to the main variables distribution.

```{r}
means_by_cluster_size <- 
  df_clustering_info_added %>% 
  group_by(analysis_info.clustering.belongs_to_small_cluster.k10) %>% 
  select_if(cols_to_select) %>% 
  summarise_all(mean) %>% 
  round(2) %>% 
  rename_with(function(name) paste("mean", name, sep = "_")) %>% 
  rename(small_cluster = 
           mean_analysis_info.clustering.belongs_to_small_cluster.k10) %>% 
  mutate(small_cluster = as.logical(small_cluster)) %>% 
  arrange(desc(mean_size))
means_by_cluster_size %>% 
  paged_table()
```

The result is amazing! The risk of the virus increase so much by belonging to a small group.

Let's express this in percentaje.

```{r}
increment_by_small_cluster <- 
  (means_by_cluster_size[1,] / 
  means_by_cluster_size[2,]) %>% 
  select(-small_cluster)
increment_by_small_cluster %>% 
  paged_table()
```

```{r}
increment_by_small_cluster_percentage <- 
  (increment_by_small_cluster - 1) %>% 
  select_if(function(x) !is.infinite(x)) %>% 
  unlist() %>% 
  percent()
increment_by_small_cluster_percentage
```

## Summarize results.

Looking at the results of the dendogram and knowing the distribution of clusters, seem that most of the viruses follows common patterns, but there are some others that breaks radically with these patterns.

Virus that don't follow patterns tend to:

- (size) Infect x8 more computers
- (community_reputation) Increase bad reputation x438
- (harmless_votes) Are the only ones with harmless votes.
- (malicious_votes) Increase the chance to have malicious votes x67
- (positives) Decrease positives by 12%
- (times_submitted) Decrease times submitted by 54%
- (total) Increase total by 3,75%
- (unique_source) Increase the unique sources x16

It is interesting to see how the most dangerous viruses:

- Are the only ones who have harmless votes.
- Request more permissions.

## Conclusion

Concluding, virus that don't follow the commons patterns tends to be much more dangerous than the others.

Signs of dangerous virus:

- Has high community reputation.
- Has harmless votes.
- Request many permissions.
- Has many unique sources.

## Save dataframe

Add the clustering information to the data frame.

```{r}
cluster_info_cols_index <- 
  df_clustering_info_added %>% 
  colnames() %>% 
  str_detect("^analysis_info") %>% 
  which()

cluster_info_cols <- 
  df_clustering_info_added %>% 
  select(all_of(cluster_info_cols_index))
```

```{r}
virusTotal_cluster <- 
  virusTotal %>% 
  select(-...1) %>% 
  cbind(cluster_info_cols)
```

```{r}
write.csv(virusTotal_cluster, path_export)
```













