---
title: "FCA"
author: "Ángel Sergio Sánchez Marín"
date: '2022-06-08'
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# FCA

```{r message=FALSE}
library(fcaR)
library(dplyr)
library(readr)
library(tidyverse)
library(rmarkdown)
```

```{r}
path_import_o = "data/original.csv"
```

```{r}
virus_o <- read_csv(path_import_o)
```

## Let's try to find out which antivirus share the same graphic engine. To do this we must study how the other antiviruses behave when one detects the files as a virus.

## we create a reduced dummie of the data frame to reduce the time it takes to find implications and concepts

```{r}

virus_o <- virus_o %>% select(contains('detected')) %>% select(c(1:30,64)) %>% 
  slice(1:60)

# we leave only the antivirus name 

colnames(virus_o)<-gsub("scans.","",colnames(virus_o))
colnames(virus_o)<-gsub(".detected","",colnames(virus_o))
```

## We leave only the antivirus name

```{r}

colnames(virus_o)<-gsub("scans.","",colnames(virus_o))
colnames(virus_o)<-gsub(".detected","",colnames(virus_o))
```

```{r eval=FALSE}
virus_o %>% 
  paged_table()
```

## Transform boolean values to character to apply nominal scaling afterwards.

## Null values symbolize that the file could not be processed.

```{r}

virus_detected <- sapply(virus_o[, c(1:31)], as.character)

# The na values correspond to data that could not be processed.
virus_detected[is.na(virus_detected)]<-"UNABLE TO PROCESS"
```

```{r eval=FALSE}
virus_detected %>% 
  paged_table()
```

## We go through the columns and apply a nominal scaling.

```{r}
#We create the formal context
fc <- FormalContext$new(virus_detected)

for (i in 1:ncol(virus_o)){
  fc$scale(colnames(virus_o[i]), type = "nominal")
}

```

## Removes duplicated attributes and objects.

```{r include=FALSE}
fc$clarify(TRUE)
```

## Implications and concepts with some examples

```{r echo=TRUE}
fc$find_implications()
```

```{r ,echo=TRUE}
fc$find_concepts()
```

```{r ,echo=TRUE}
fc$implications[77]
```

```{r ,echo=TRUE}
fc$concepts[16]
```

## Avast acquired AVG in 2016, so they currently use the same engine with some different features. If our study has been carried out correctly, AVG=TRUE (detected as virus) should be a closure to Avast = TRUE.

```{r closure}
S <- Set$new(attributes = fc$attributes)
S$assign("Avast = TRUE" = 1)
cat("Given the attribute:")
S
cat("It has as closed::")
# Compute the intent of S
fc$closure(S)
```

## Even with a reduced data frame, this assumption holds true.

## Our study could also be used to look at the effectiveness of an antivirus. As long as we compare it with an antivirus that we consider reliable.
