---
title: "Virustotal_aRules"
author: "Ángel Sergio Sánchez Marín"
date: '2022-05-18'
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r librerias}
library(dplyr)
library(arules)
library(magrittr)
library(rmarkdown)
library(readr)
library(stringr)
library(tidyverse)
```

```{r}
path_import = "../virusTotal/data/original.csv"
path_export = "../virusTotal/data/dataDiscretized.csv"
```

### Import of the VirusTotal dataset.

```{r import, include=FALSE}
virus <- read_csv(path_import)
```

*In this exercise we are going to observe the access to permissions by files together with the risk they present.*

*As there are a large number of viruses each with different objectives, the association rules generated can be used to predict which permissions a particular virus wants to access along with the probability that it will do so.*

*We have also added the number of positives to see the risk factor of the different permit accesses.*

### We replace all the NA in the values of the columns permits, since we are going to relate them later with the numbers of positives.

```{r read}

 

replace_na_which_colname_match <- 
  function(df, pattern, replacement){
    cols_to_replace <- 
      df %>% 
      colnames() %>% 
      str_which(pattern)
    df_replaced_cols <- 
      df %>% 
      select(all_of(cols_to_replace)) %>% 
      sapply(function(col) replace_na(col, replacement))
    df_without_replaced_cols <- 
      df %>% 
      select(-all_of(cols_to_replace))
    return(cbind(df_without_replaced_cols, df_replaced_cols))
  }

pattern <- "additional_info.androguard.RiskIndicator.PERM"

virus <- 
  virus %>% 
  replace_na_which_colname_match(pattern, 0)




```

### We select the rules.

```{r select}
df_rules <- virus %>%
   select(c( positives,
             additional_info.androguard.RiskIndicator.PERM.DANGEROUS,
             additional_info.androguard.RiskIndicator.PERM.GPS,
             additional_info.androguard.RiskIndicator.PERM.NORMAL,
             additional_info.androguard.RiskIndicator.PERM.PRIVACY,
             additional_info.androguard.RiskIndicator.PERM.INTERNET,
             additional_info.androguard.RiskIndicator.PERM.INSTANT,
             additional_info.androguard.RiskIndicator.PERM.SMS,
             additional_info.androguard.RiskIndicator.PERM.MONEY,
             additional_info.androguard.RiskIndicator.PERM.SYSTEM,
             additional_info.androguard.RiskIndicator.PERM.CALL,
             additional_info.androguard.RiskIndicator.PERM.SIGNATURE,
             additional_info.androguard.RiskIndicator.PERM.DEVELOPMENT,
             additional_info.androguard.RiskIndicator.PERM.PRE23,
             additional_info.androguard.RiskIndicator.PERM.APPOP,
             additional_info.androguard.RiskIndicator.PERM.EPHEMERAL,
             additional_info.androguard.RiskIndicator.PERM.PREINSTALLED,
             additional_info.androguard.RiskIndicator.PERM.INSTALLER,
             additional_info.androguard.RiskIndicator.PERM.RUNTIME
             ))

colnames(df_rules)<-gsub("additional_info.androguard.RiskIndicator.PERM.","",
                         colnames(df_rules))
View(df_rules)
```

```{r view, include=FALSE}
View(df_rules)
```

### We will discretize the values using quantiles.

### The function does not apply to columns with few values, as they do not provide much information.

```{r discretize}

labels <- 
  function(n){
    if(n == 5){
      return(c("very low", "low", "medium", "high", "very high"))
    }else if(n == 4){
      return(c("very low", "low", "high", "very high"))
    }else if(n == 3){
      return(c("low", "medium", "high"))
    }else if(n == 2){
      return(c("low", "high"))
    }else{
      stop("Not avalible")
    }
  }

cut_by_quantiles <- 
  function(col){
    quantiles <- 
      col %>% 
      na.omit() %>% 
      quantile() %>% 
      unique()
    if(length(quantiles) > 2){
      col <- 
        col %>% 
        cut(breaks = quantiles, labels = labels(length(quantiles)-1), include.lowest = TRUE)
    }
    return(col)
  }

df_cut_by_quantiles <- 
  function(df){
    df_without_numeric <- 
      df[sapply(df, function(col) !is.numeric(col))]
    df_numeric <- 
      df %>% 
      select_if(is.numeric)
    df_numeric <- 
      df_numeric %>% 
      lapply(cut_by_quantiles)
    return(cbind(df_without_numeric, df_numeric))
  }

df_test <- 
  df_rules  %>% 
  df_cut_by_quantiles()
View(df_test)
```

### We eliminate the untouched columns.

```{r}
df_discretize <- df_test %>% select(c( 1,2,4,5,8,9))
View(df_discretize )
```

```{r}
View(df_discretize )
```

```{r}
write.csv(df_discretize, path_export)
```

### We use the apriori method with standard confidence and support values.

```{r apriori}
rules <- apriori(df_discretize, parameter = list(supp=0.1, target="rules", conf=0.8))

```

### We remove redundant rules.

```{r}
redundantes <- is.redundant(rules)
r_no_redundante <- rules[!redundantes]
inspect(head(r_no_redundante))

significantes <- is.significant(r_no_redundante)
r_significantes <- r_no_redundante[significantes]
inspect(head(r_significantes))

```

### Display

```{r}
library(arulesViz)
subrules <- subset(r_significantes, lift>3)

plot(subrules, method="graph", engine="htmlwidget", 
     igraphLayout = "layout_in_circle")
```

We see that if the number of positives is low, the danger is medium, indicating that the antivirus used by VirusTotal is of good quality.

Another curious relationship is between privacy and money, which show a high dependency. Because to access your money they must first know your bank account details. There is also a similar relationship between sms and privacy.
