---
title: "From JSON to CSV"
author: "Pedro Antonio Benito Rojano"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# From JSON to CSV

We have a dataset in JSONs files, but the suitable type of file is CSV. So we need to convert the JSONs to a CSV file. 

Steps:

1. Load the JSONs files.
2. Convert JSONs to dataset.
3. Save dasaset as CSV file.

## Configuration

Set the path to import the JSONs files.

```{r warning=FALSE}
path_import_jsons <- "F:/Laboratorio de Computación/Android"
```

Set the path to export the dataset as CSV.

```{r warning=FALSE}
path_export_csv <- "C:/Users/angel/Desktop/R_TRABAJO/VirusTotal-DataMining/VirusTotal/virusTotal.csv"
```

## Libraries

```{r message=FALSE}
library(dplyr)
library(tidyjson)
```

## Convertion

### Load JSON

#### Define functions

Define function for list a forlder.

```{r}
list_folder <- 
  function(path){
    list.files(path = path) %>% 
    sapply(function(json) paste(path, json, sep="/"))
  }
```

Define function for transform JSON to a tibble.

```{r}
json_to_df <- 
  function(path){
    path %>% 
      read_json() %>% 
      spread_all() %>% 
      tibble()
  }
```

Define function for that generate a dataset from a folder with JSONs files.

```{r}
jsons_folder_to_df <- 
  function(path){
    path %>% 
      list_folder() %>% 
      lapply(json_to_df) %>% 
      bind_rows() %>% 
      lapply(as.character)
  }
```

#### Load with functions

```{r}
View(df)
df <- jsons_folder_to_df(path_import_jsons)
```

### Save as CSV

```{r}
write.csv(df, path_export_csv)
```


```{r}
virus <- read.csv('virusTotal.csv')
View(virus)
virus$perm
```


