---
title: "From JSON to CSV"
author: "Pedro Antonio Benito Rojano"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# From JSON to CSV

We have a dataset in JSONs files, but the suitable type of file is CSV. So we need to convert the JSONs to a CSV file. 

Steps:

1. Load the JSONs files.
2. Convert JSONs to dataset.
3. Save dasaset as CSV file.

## Configuration

Set the path to import the JSONs files.

```{r}
path_import_jsons <- "data/json"
```

Set the path to export the dataset as CSV.

```{r}
path_export_csv <- "data/virusTotal.csv"
```

## Libraries

Library for pipe working.

```{r message=FALSE}
library(dplyr)
```

Library for manipulate JSONs.

```{r}
library(tidyjson)
```

Libraries for parallel execution.

```{r warning=FALSE}
library(parallel)
```

## Defining functions

### Auxiliary functions

Define function for list a forlder.

```{r}
list_folder <- 
  function(path){
    list.files(path = path) %>% 
    sapply(function(json) paste(path, json, sep="/"))
  }
```

Define function for transform JSON to a dataset

```{r}
json_to_df <- 
  function(path){
    path %>% 
      read_json() %>% 
      spread_all() %>% 
      tibble()
  }
```

Define function for generate a dataset from a folder with JSONs files.

```{r}
jsons_folder_to_df <- 
  function(path){
    path %>% 
      list_folder() %>% 
      lapply(json_to_df) %>% 
      bind_rows()
  }
```

Define function for save a data frame to CSV.

```{r}
save_df_to_csv <- 
  function(df, path){
    df %>% 
      lapply(as.character) %>% 
      write.csv(path_export_csv)
  }
```

Define function for generate a CSV datset from a JSON folder.

```{r}
from_json_to_csv <- 
  function(json_folder_path, csv_path){
    jsons_folder_to_df(json_folder_path) %>% 
      save_df_to_csv(csv_path)
  }
```

### Parallel

Define function for setup the cluster, with needed libraries

```{r}
setup_cluster <- 
  function() {
    cluster <- 
      detectCores() %>% 
      makeCluster()
    clusterEvalQ(cluster, expr = library(tidyjson))
    clusterEvalQ(cluster, expr = library(dplyr))
    return(cluster)
  }
```

Define function for generate a dataset from a folder with JSONs files by parallel execution.

```{r}
jsons_folder_to_df_parallel <- 
  function(cluster, path){
    json_paths <- list_folder(path)
    df <- 
      cluster %>% 
      parLapply(json_paths, json_to_df) %>% 
      bind_rows() 
    json_paths
    return(df)
  }
```

Define function for generate a CSV datset from a JSON folder by parallel execution.

```{r}
from_json_to_csv_parallel <- 
  function(json_folder_path, csv_path){
    cluster <- setup_cluster() 
    cluster %>% 
      jsons_folder_to_df_parallel(json_folder_path) %>% 
      save_df_to_csv(csv_path)
    stopCluster(cluster)
  }
```

## From JSON to CSV

Using defined functions for converting from JSON to CSV.

```{r}
from_json_to_csv_parallel(path_import_jsons, path_export_csv)
```


